Openness and Computational Reproducibility in plant pathology: where do we stand and a way forward

### 2022-06-02 analysis/paper/paper.Rmd
Insert missing \newpage in figure section
Revise figures

load(file = here("inst/extdata/m_f.Rda"))
load(file = here("inst/extdata/m_g.Rda"))
load(file = here("inst/extdata/m_h.Rda"))
load(file = here("inst/extdata/m_i.Rda"))
load(file = here("inst/extdata/m_j.Rda"))
load(file = here("inst/extdata/m_k.Rda"))

pp_m_f <- pp_check(m_f)
pp_m_g <- pp_check(m_g)
pp_m_h <- pp_check(m_h)
pp_m_i <- pp_check(m_i)
pp_m_j <- pp_check(m_j)
pp_m_k <- pp_check(m_k)

# put them all together
p <- (pp_m_f | pp_m_g) / (pp_m_h | pp_m_i) / (pp_m_j | pp_m_k)

p <- p + plot_annotation(tag_levels = "A",
                         title = "Bayesian posterior distributions of evaluation score models")

@@

```{r irCriteria, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Evaluator agreement for five articles scored for (A) software being properly cited, (B) software used being freely and openly available, (C) computational methods used, e.g. scripts or other self-written programs were freely and openly available and (D) data represented in the paper were freely and openly available. Note that one evaluator rated an article with computational methods available as 'NA', which is reflected in that article having only four evaluations for this criteria.", cache=FALSE, fig.env='figure'}
a <- ggplot(ir,
            aes(y = doi,
                x = software_cite)) +
  ylab("DOI") +
  xlab("") +
  xlim(c(0, 3)) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  geom_density_ridges() +
  ggtitle("Software Cited")

b <- ggplot(ir,
            aes(y = doi,
                x = software_avail)) +
  ylab("") +
  xlab("") +
  geom_density_ridges() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  ) +
  xlim(c(0, 3)) +
  ggtitle("Software Available")

c <- ggplot(ir,
            aes(y = doi,
                x = comp_mthds_avail)) +
  geom_density_ridges() +
  ylab("DOI") +
  xlab("Score") +
  xlim(c(0, 3)) +
  ggtitle("Comp. Methods Available")

d <- ggplot(ir,
            aes(y = doi,
                x = data_avail)) +
  ylab("") +
  xlab("Score") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  xlim(c(0, 3)) +
  geom_density_ridges() +
  ggtitle("Data Available")

p <- (a | b) / (c | d)
p <- p + plot_annotation(tag_levels = "A",
                         title = "Evaluator agreement in article reproducibility criteria scoring")

@@

```{r irReproScore, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Overall reproducibility scores of five articles scored for evaluator agreement. Where 0 was not reproducible in any fashion and a perfect score would be 100% as evaluated for four areas software citations, software availability, computational methods availability and data availability on a 0 to 3 scale and converted to a percent of total possible score value.", cache=FALSE, fig.env='figure'}
p <- ggplot(ir,
            aes(y = doi,
                x = reproducibility_score)) +
  geom_density_ridges() +
  ggtitle("Inter-rater reproducibility score for five articles") +
  ylab("DOI") +
  xlab("Reproducibility Score (%)") +
  xlim(c(0, 100))

@@

```{r reproScore, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Reproducibility scores of 450 papers published in 21 plant pathology journals or plant pathology focused articles from other specialised journals. Where 0 was not reproducible in any fashion and a perfect score would be 100% as evaluated for four areas software citations, software availability, computational methods availability and data availability on a 0 to 3 scale and converted to a percent of total possible score value.", cache=FALSE, fig.env='figure'}
p <-
  ggplot(data = rrpp, aes(x = reproducibility_score)) +
  geom_histogram(binwidth = 10,
                 fill = "black",
                 colour = "black") +
  xlab("Reproducibility score (%)")

ggsave(
  filename = here("analysis/figures/Sparks_et_al_Figure_5.eps"),
  plot = p,
  width = 6,
  height = 6,
  units = "in",
  device = cairo_ps
)
embed_fonts(
  file = here("analysis/figures/Sparks_et_al_Figure_5.eps"),
  outfile = here("analysis/figures/Sparks_et_al_Figure_5.eps"),
  options = "-dEPSCrop"
)
include_graphics(path = here("analysis/figures/Sparks_et_al_Figure_5.eps"))
```

```{r articleScores, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Aggregated article scores for each of the four categories, (A) shows computational methods availability 0 - not available or not mentioned to 3 - freely available online to anonymous users. (B) presents 'Software availability', where '0' was 'Not available or not mentioned in the publication'; '1' was 'Expensive proprietary software that only institutions would typically purchase'; '2' was 'Used proprietary software that most individuals can afford (e.g., Excel) or SAS that is now available free for use with a university login'; and '3' was 'Used entirely open source and free software (e.g., R, Julia, Python)'; 'NA' indicates that no specialised software was necessary or used. (C) displays 'Software citated', where '0' was 'Not available or not mentioned in the publication'; '1' was 'Software mentioned by name only'; '2' was 'Software cited with version number'; and '3' was 'All software components (SAS PROCs, R, Julia or Python packages, etc.) cited'; 'NA' indicates that no specialised software was necessary or used. (D) shows 'Data availability', where '0' was 'Not available or not mentioned in the publication'; '1' was 'Available upon request to the author; '2' was 'Online, but inconvenient or non-permanent (e.g., login needed, paywall, FTP server, personal lab website that may disappear, or may have already disappeared)'; and '3' was 'Freely available online to anonymous users for foreseeable future (e.g., archived using Zenodo, dataverse or university library or some other proper archiving system)'; 'NA' indicates that no data were generated, e.g., a methods paper.", cache=FALSE, fig.env='figure'}

rrpp_long <-
  pivot_longer(rrpp,
               cols = c(
                 "comp_mthds_avail",
                 "software_avail",
                 "software_cite",
                 "data_avail"
               ))

p <- ggplot(rrpp_long, aes(x = as.factor(value))) +
  geom_bar(fill = "black") +
  ylab("Count") +
  xlab("Article Score") +
  ggtitle("Reproducibility criteria index values") +
  facet_wrap(name ~ .)
p <- tag_facet(
  p = p,
  tag_pool = LETTERS,
  x = 5,
  family = "Arial",
  open = "",
  close = ""
)

ggsave(
  filename = here("analysis/figures/Sparks_et_al_Figure_6.eps"),
  plot = p,
  width = 6,
  height = 6,
  units = "in",
  device = cairo_ps
)
embed_fonts(
  file = here("analysis/figures/Sparks_et_al_Figure_6.eps"),
  outfile = here("analysis/figures/Sparks_et_al_Figure_6.eps"),
  options = "-dEPSCrop"
)
include_graphics(path = here("analysis/figures/Sparks_et_al_Figure_6.eps"))
```

@@

rsvg_ps(svg = charToRaw(p),
        file = here("analysis/figures/Sparks_et_al_Figure_7.eps"))
embed_fonts(
  file = here("analysis/figures/Sparks_et_al_Figure_7.eps"),
  outfile = here("analysis/figures/Sparks_et_al_Figure_7.eps"),
  options = "-dEPSCrop"
)

include_graphics(path = here("analysis/figures/Sparks_et_al_Figure_7.eps"))
```

>>>>

load(file = here("inst/extdata/m_f1.Rda"))
load(file = here("inst/extdata/m_f2.Rda"))
load(file = here("inst/extdata/m_f3.Rda"))
load(file = here("inst/extdata/m_g1.Rda"))
load(file = here("inst/extdata/m_g2.Rda"))
load(file = here("inst/extdata/m_g3.Rda"))
load(file = here("inst/extdata/m_h1.Rda"))
load(file = here("inst/extdata/m_h2.Rda"))
load(file = here("inst/extdata/m_h3.Rda"))
load(file = here("inst/extdata/m_i1.Rda"))
load(file = here("inst/extdata/m_i2.Rda"))
load(file = here("inst/extdata/m_i3.Rda"))
load(file = here("inst/extdata/m_j1.Rda"))
load(file = here("inst/extdata/m_j2.Rda"))
load(file = here("inst/extdata/m_j3.Rda"))

mod_list <- list(
  pp_check(m_f1) +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_blank()
    ),
  pp_check(m_f2) +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_blank()
    ),
  pp_check(m_f3) +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_blank()
    ),
  pp_check(m_g1) +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_blank()
    ),
  pp_check(m_g2) +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_blank()
    ),
  pp_check(m_g3) +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_blank()
    ),
  pp_check(m_h1) +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_blank()
    ),
  pp_check(m_h2) +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_blank()
    ),
  pp_check(m_h3) +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_blank()
    ),
  pp_check(m_i1),
  pp_check(m_i2),
  pp_check(m_i3)
)

p <- wrap_plots(mod_list, nrow = 4) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A",
                  title = "Bayesian posterior distributions of evaluation score models")

@@

```{r irCriteria, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Evaluator agreement for five articles scored for (A) computational methods (code) used, e.g. scripts or other self-written programs were freely and openly available and (B) data represented in the paper were freely and openly available. Note that one evaluator rated an article with computational methods available as 'NA', which is reflected in that article having only four evaluations for this criteria.", cache=FALSE, fig.env='figure'}
a <- ggplot(ir,
            aes(y = doi,
                x = comp_mthds_avail)) +
  geom_density_ridges() +
  ylab("DOI") +
  xlab("Score") +
  xlim(c(0, 3)) +
  ggtitle("Code Available")

b <- ggplot(ir,
            aes(y = doi,
                x = data_avail)) +
  ylab("") +
  xlab("Score") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  xlim(c(0, 3)) +
  geom_density_ridges() +
  ggtitle("Data Available")

p <- a + b
p <- p + plot_annotation(tag_levels = "A",
                         title = "Evaluator agreement in article reproducibility criteria scoring")

@@

```{r articleScores, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Aggregated article scores for each of the three categories evaluated, (A) presents 'Software availability', where '0' was 'Not available or not mentioned in the publication'; '1' was 'Expensive proprietary software that only institutions would typically purchase'; '2' was 'Used proprietary software that most individuals can afford (e.g., Excel) or SAS that is now available free for use with a university login'; and '3' was 'Used entirely open source and free software (e.g., R, Julia, Python)'; 'NA' indicates that no specialised software was necessary or used. (B) displays 'Code availability', where '0' was 'Not available or not mentioned in the publication'; '1' was 'Available upon request to the author; '2' was 'Online, but inconvenient or non-permanent (e.g., login needed, paywall, FTP server, personal lab website that may disappear, or may have already disappeared)'; and '3' was 'Freely available online to anonymous users for foreseeable future (e.g., archived using Zenodo, dataverse or university library or some other proper archiving system)'; 'NA' indicates that no code was created to conduct the work that was detectable. (C) shows 'Data availability', where '0' was 'Not available or not mentioned in the publication'; '1' was 'Available upon request to the author; '2' was 'Online, but inconvenient or non-permanent (e.g., login needed, paywall, FTP server, personal lab website that may disappear, or may have already disappeared)'; and '3' was 'Freely available online to anonymous users for foreseeable future (e.g., archived using Zenodo, dataverse or university library or some other proper archiving system)'; 'NA' indicates that no data were generated, e.g., a methods paper.", cache=FALSE, fig.env='figure'}

a <- ggplot(rrpp, aes(x = as.factor(software_avail))) +
  geom_bar(fill = "black") +
  ylab("Count") +
  xlab("Article Score") +
  ggtitle("Software")

b <- ggplot(rrpp, aes(x = as.factor(comp_mthds_avail))) +
  geom_bar(fill = "black") +
  ylab("Count") +
  xlab("Article Score") +
  ggtitle("Code")

c <- ggplot(rrpp, aes(x = as.factor(data_avail))) +
  geom_bar(fill = "black") +
  ylab("Count") +
  xlab("Article Score") +
  ggtitle("Data")

p <- a + b + c
p <- p +
  plot_annotation(tag_levels = "A")

@@

@@

rsvg_ps(svg = charToRaw(p),
        file = here("analysis/figures/Sparks_et_al_Figure_5.eps"))
embed_fonts(
  file = here("analysis/figures/Sparks_et_al_Figure_5.eps"),
  outfile = here("analysis/figures/Sparks_et_al_Figure_5.eps"),
  options = "-dEPSCrop"
)

include_graphics(path = here("analysis/figures/Sparks_et_al_Figure_5.eps"))
```
